\documentclass[10pt,conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\hypersetup{colorlinks=true,urlcolor=black,linkcolor=black,citecolor=black}
% \usepackage{subfig}
\renewcommand{\sectionautorefname}{Section}
\usepackage{enumitem}
\usepackage{balance}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{booktabs}
% \usepackage{adjustbox}
\usepackage{multirow}
\usepackage{comment}
\usepackage{threeparttable}
\usepackage{microtype}
\usepackage{relsize,xspace}
\usepackage{cite}
\usepackage{textcomp}
\usepackage{mdframed}
\usepackage{framed}
\usepackage{enumitem}
\usepackage{MnSymbol,wasysym}
\usepackage{subfigure}
\usepackage{mdframed}
\usepackage{framed}
\usepackage{float}
\usepackage[normalem]{ulem} % [normalem] prevents the package from changing the default behavior of `\emph` to underline.


% \usepackage[dvipsnames]{xcolor}


\usepackage{listings}
\usepackage{caption}

%\usepackage{minted}
\usepackage{color, soul}

\definecolor{dark_green)}{rgb}{0.0, 0.5, 0.0}

\newcommand{\tm}[1]{{\color{blue}\textsf{TM}[\smaller\sffamily #1}]}
\newcommand{\az}[1]{{\color{orange}\textsf{AZ}[\smaller\sffamily #1}]}
\newcommand{\jb}[1]{{\color{red}\textsf{JB}[\smaller\sffamily #1}]}
\newcommand{\sd}[1]{{\color{magenta}\textsf{SD}[\smaller\sffamily\color{magenta} #1}]}
\newcommand{\ad}[1]{{\color{violet}\textsf{AD}[\smaller\sffamily #1}]}
\newcommand{\cd}[1]{{\color{cyan}\textsf{CR}[\smaller\sffamily #1}]}

\newcommand{\java}{\textsf{Java}\xspace}
\newcommand{\cp}{\textsf{C}\xspace}
\newcommand{\cpp}{\textsf{C++}\xspace}
%\newcommand{\java}{\textsf{Java}\xspace}
\newcommand{\py}{\textsf{Python}\xspace}
\newcommand{\rb}{\textsf{Ruby}\xspace}
\newcommand{\go}{\textsf{Go}\xspace}
\newcommand{\librariesio}{\textsf{Libraries.io}\xspace}

\newcommand{\git}{\textit{git}\xspace}
\newcommand{\rqOne}{$RQ_1$ Do the variant forks and their upstream counterparts have common maintainers?}
\newcommand{\rqOneOne}{$RQ_{1.1}$ \emph{How many of the original developers of the mainline maintained the variant in its first 6 months?}}
\newcommand{\rqOneTwo}{$RQ_{1.2}$ \emph{Do the variant and mainline have common active maintainers?}}
\newcommand{\rqTwo}{$RQ_{2}$ What was the motivation for creating the variant fork of an upstream project?}
\newcommand{\rqTwoOne}{$RQ_{2.1}$ \emph{Was the motivation for creating the variant an individual or a community decision?}}
\newcommand{\rqTwoTwo}{$RQ_{2.2}$ \emph{What was the motivation for creating the variant of the mainline project?}}
\newcommand{\rqTwoThree}{$RQ_{2.3}$ \emph{What are the motivation details relating to the motivation in $RQ_{2.2}$?}}
\newcommand{\rqThree}{$RQ_{3}$ Do the upstream and variant fork still interact?}
\newcommand{\rqThreeOne}{$RQ_{3.1}$ \emph{Do the variant forks and the upstream still discuss the main directions of the project?}}
\newcommand{\rqThreeTwo}{$RQ_{3.2}$ \emph{Do the variant developers integrate changes to and from the upstream repository?}}

\newcommand{\gh}{GitHub\xspace}
\newcommand{\ra}{\ensuremath{\rightarrow}}
\newcommand{\nd}{\vspace{1mm}\noindent}

\makeatletter
\def\endthebibliography{%
  \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
  \endlist
}
\makeatother



\begin{document}

\title{Motivations Behind Variant Forks Creation and Maintenance}

\title{Variant Forks -- Motivations and Impediments}

% \author{\IEEEauthorblockN{John Businge,\IEEEauthorrefmark{1}
% 		Alexandre Decan,\IEEEauthorrefmark{2}
% 	Ahmed Zerouali,\IEEEauthorrefmark{3}
% 	Tom Mens\IEEEauthorrefmark{2} and
% 	Serge Demeyer,\IEEEauthorrefmark{1}}
% 	\IEEEauthorblockA{\IEEEauthorrefmark{1}University of Antwerp, Antwerp, Belgium}
% 	\IEEEauthorblockA{\IEEEauthorrefmark{2}University of Mons, Mons, Belgium}
% 	\IEEEauthorblockA{\IEEEauthorrefmark{3} Vrije Universiteit Brussels, Brussels, Belgium}
% 	}


\author{\IEEEauthorblockN{Author 1,\IEEEauthorrefmark{1}
 		Author 2,\IEEEauthorrefmark{2}
 		Author 3,\IEEEauthorrefmark{3}
 		Author 4\IEEEauthorrefmark{2} and
 		Author 5,\IEEEauthorrefmark{1}}
 	\IEEEauthorblockA{\IEEEauthorrefmark{1}Anonymous Department 1}
 	\IEEEauthorblockA{\IEEEauthorrefmark{2}Anonymous Department 1}
 	\IEEEauthorblockA{\IEEEauthorrefmark{3}Anonymous Department 1}
 	}


\maketitle

\begin{abstract}
Some developers are reusing existing software software (OSS) hosted on social coding platforms by splitting-off a new development branch, often to steer the development into another direction, without the intention to  contribute back. We call these kind of forks variants of the original (mainline) projects.
We conduct an exploratory study on variant forks to understand the motivations behind variant fork creation and maintenance. Our study is based on a survey conducted on 112 maintainers of 105 active OSS variants projects hosted on GitHub a very popular social coding platform. Our study confirms and extends the findings of previous studies. Among the findings, we identify a number of fine-grained motivations behind the creation and maintenance of variants. We also identify concrete reasons why there is limited interaction between the mainline and variant projects. Based on our findings, we discuss future research direction in the line of efficient software reuse between the variants and the mainline counterparts.
\
\sd{-- Alternative abstract--}

Social coding platforms centred around git (i.e. \gh, GitLab, Bitbucket) provide explicit facilities to share code between projects: forks, pull requests, cherry-picking to name but a few.
Variant forks are an interesting phenomenon in that respect, as it permits for different projects to peacefully co-exist, yet explicitly acknowledge the common ancestry.
Several researchers analysed forking practices on open source platforms and observed that variant forks get created frequently.
However, today little is known on the motivations for launching such a variant fork.
Is it mainly technical (e.g. diverging features), governance (e.g. diverging communities), legal (e.g. diverging licences) or do other factors come into play.
In this paper we report the results of an exploratory study on the motivations concerning variant fork creation and maintenance.
We conducted a survey among 112 maintainers of 105 active OSS variants projects hosted on \gh a very popular social coding platform.
Our study extends previous findings, identifying a number of common motivations for launching a variant fork and listing concrete impediments for maintaining the co-existing projects.
end{abstract}

\begin{IEEEkeywords}
Mainlines, Variants, GitHub, Software ecosystems, Maintenance, Variability
\end{IEEEkeywords}

\input{1intro}
\input{2background_and_related_work}
\input{3research_questions}
\input{4methods}
\input{5results_and_discussion}
\input{6threats_and_conclusion}


%TOM: FORCE NEW PAGE FOR REFERENCES (since we have 10+1page)
\newpage
\tm{John, clean up references, for example, make sure that all titles of papers properly use capitalisation. There are many places where this is not the case because by default latex puts all words in titles of a citation in lowercase. So you have to force words that you need to uppercase by putting them between curly braces.}
\tm{There are other things that can be reduced, e.g. no need to put "Proceedings of", no need to put the conference number (e.g. 19th) or conference year as part of the conference title. Otherwise the year will be shown twice. Be consistent in using DOI or URL for citations. Either use it everywhere or nowhere. Preferably consistently use the DOI instead of some publisher-specific URL if the DOI is available.}

\bibliographystyle{IEEEtran}
\typeout{}
\bibliography{biblio}

\end{document}
