\documentclass[10pt,conference]{IEEEtran}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{hyperref}
\hypersetup{colorlinks=true,urlcolor=black,linkcolor=black,citecolor=black}
% \usepackage{subfig}
\renewcommand{\sectionautorefname}{Section}
\usepackage{enumitem}
\usepackage{balance}
\usepackage{xcolor}
\usepackage{graphicx}
\usepackage{booktabs}
% \usepackage{adjustbox}
\usepackage{multirow}
\usepackage{comment}
\usepackage{threeparttable}
\usepackage{microtype}
\usepackage{relsize,xspace}
\usepackage{cite}
\usepackage{textcomp}
\usepackage{mdframed}
\usepackage{framed}
\usepackage{enumitem}
\usepackage{MnSymbol,wasysym}
\usepackage{subfigure}
\usepackage{mdframed}
\usepackage{framed}
\usepackage{float}
\usepackage[normalem]{ulem} % [normalem] prevents the package from changing the default behavior of `\emph` to underline.


% \usepackage[dvipsnames]{xcolor}


\usepackage{listings}
\usepackage{caption}

%\usepackage{minted}
\usepackage{color, soul}

\definecolor{dark_green)}{rgb}{0.0, 0.5, 0.0}

\newcommand{\tm}[1]{{\color{blue}\textsf{TM}[\smaller\sffamily #1}]}
\newcommand{\az}[1]{{\color{orange}\textsf{AZ}[\smaller\sffamily #1}]}
\newcommand{\jb}[1]{{\color{red}\textsf{JB}[\smaller\sffamily #1}]}
\newcommand{\sd}[1]{{\color{magenta}\textsf{SD}[\smaller\sffamily\color{magenta} #1}]}
\newcommand{\ad}[1]{{\color{violet}\textsf{AD}[\smaller\sffamily #1}]}
\newcommand{\cd}[1]{{\color{cyan}\textsf{CR}[\smaller\sffamily #1}]}

\newcommand{\java}{\textsf{Java}\xspace}
\newcommand{\cp}{\textsf{C}\xspace}
\newcommand{\cpp}{\textsf{C++}\xspace}
%\newcommand{\java}{\textsf{Java}\xspace}
\newcommand{\py}{\textsf{Python}\xspace}
\newcommand{\rb}{\textsf{Ruby}\xspace}
\newcommand{\go}{\textsf{Go}\xspace}
\newcommand{\librariesio}{\textsf{Libraries.io}\xspace}
\newcommand{\gh}{GitHub\xspace}
\newcommand{\sourceforge}{SourceForge\xspace}
\newcommand{\git}{\textit{git}\xspace}



\newcommand{\rqOne}{$RQ_1$ Why do developers create and maintain variants of a (mainline) original projects?}
\newcommand{\rqOneOne}{$RQ_{1.1}$ \textbf{Was the motivation for creating the variant an individual or a community decision?}}
\newcommand{\rqOneTwo}{$RQ_{1.2}$ \textbf{What was the motivation for creating the variant of the mainline project?}}
\newcommand{\rqOneThree}{$RQ_{1.3}$ \textbf{What are the motivation details relating to the motivation in $RQ_{1.2}$?}}


% \newcommand{\rqThree}{$RQ_{3}$ Do the upstream and variant fork still interact?}


\newcommand{\rqTwo}{$RQ_{2}$ How do variant projects evolve w.r.t the mainline?}
\newcommand{\rqTwoOne}{$RQ_{2.1}$ \emph{How many of the original developers of the mainline maintained the variant in its first 6 months?}}
\newcommand{\rqTwoTwo}{$RQ_{2.2}$ \emph{Do the variant and mainline have common active maintainers?}}
\newcommand{\rqTwoThree}{$RQ_{2.3}$ \emph{Do the variant forks and the upstream still discuss the main directions of the project?}}
\newcommand{\rqTwoFour}{$RQ_{2.4}$ \emph{Do the variant developers integrate changes to and from the upstream repository?}}

\newcommand{\ra}{\ensuremath{\rightarrow}}
\newcommand{\nd}{\vspace{1mm}\noindent}

\makeatletter
\def\endthebibliography{%
  \def\@noitemerr{\@latex@warning{Empty `thebibliography' environment}}%
  \endlist
}
\makeatother